---
title: "Infrastructure and Citation Analysis"
author: "Awan Afiaz - Graduate Student"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: yes
    code_download: true 
    code_folding: 'hide'
    theme: cosmo
    highlight: pygments
---

```{r setup_impact, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,  fig.align='center')
## Required packages
library(tidyverse)
library(magrittr)
library(DT)
library(scales)
library(patchwork) ## plot side by side
library(here)
```


## Part I: Data loading and check

```{r loadData_impact, echo=TRUE}
## Load data file1: The Software Impact Dataset from John
wdat <- read.table(file='itcr_softwarekg_articles.tsv', sep = '\t', header = TRUE, fill = TRUE) %>% 
  rename(toolName = name_from_itcr_simplified) %>% 
  mutate(toolName = str_trim(toolName))

## Tool Analysis dataset, modified version with all the edited/recoded variables
tdat <- read_csv(file = "tdat_edited.csv", show_col_types = FALSE)  %>% 
  rename(toolName = `tool name`) %>% 
  mutate(toolName = str_trim(toolName))
```


```{r}
 
health <- read_csv(here::here("Tool_data_health_metrics.csv"))

count(health, health_metrics)
health %<>% mutate(simple_health_metrics = case_when(
  str_detect(health_metrics, 
"no|No|unknown status|no badge|Not for core software|CI but not badge|\\?") ~ "No",
 str_detect(health_metrics, 
"Yes|yes|Test badges|CI and coverage badges|bioconductor only| with badge")~ "Yes",
  TRUE ~ health_metrics
))
 
identical(tdat$toolName, health$`tool name`)
 
tdat$simple_health_metrics <- health$simple_health_metrics 
```

## 1. Proportion of softwares with health metrics shown
```{r, fig.align='center'}
tdat %>% 
    mutate(simple_health_metrics = case_when(is.na(simple_health_metrics)~ "NA", TRUE ~simple_health_metrics))%>%
  ggplot(aes(x = simple_health_metrics, fill = simple_health_metrics)) +
  geom_bar( width = 0.5) + 
  geom_text(stat = 'count', aes(label = paste0(after_stat(round(100*count/sum(count),1)), "%")), vjust = 1.5) +
  labs(x = "softwares with DOI", y = "Frequency", caption = "Distribution of softwares with health metrics") +
  scale_y_continuous(n.breaks = 10) + 
  scale_fill_brewer(palette = "Set2", direction = -1) +
  theme_light() 

tdat %>% 
    drop_na(simple_health_metrics) %>%
  ggplot(aes(x = simple_health_metrics, fill = simple_health_metrics)) +
  geom_bar( width = 0.5) + 
  geom_text(stat = 'count', aes(label = paste0(after_stat(round(100*count/sum(count),1)), "%")), vjust = 1.5) +
  labs(x = "softwares with DOI", y = "Frequency", caption = "Distribution of softwares with health metrics") +
  scale_y_continuous(n.breaks = 10) + 
  scale_fill_brewer(palette = "Set2", direction = -1) +
  theme_light() 

tdat %>%
  mutate(simple_health_metrics = case_when(is.na(simple_health_metrics)~ "NA", TRUE ~simple_health_metrics))%>%
  ggplot(aes(x = simple_health_metrics, fill = simple_health_metrics)) +
  geom_bar( width = 0.5) + 
  geom_text(stat = 'count', aes(label = after_stat(count), vjust = 1.5)) +
  facet_wrap(~`class/type`) +
  labs(x = "softwares with health metrics shown", y = "Frequency", caption = "Distribution of softwares with health metrics shown") +
  scale_y_continuous(n.breaks = 10) + 
  scale_fill_brewer(palette = "Set2", direction = -1) +
  theme_light() +
  theme(strip.background =element_rect(fill="steelblue3"),
        strip.text = element_text(colour = 'white'), 
        panel.grid = element_blank())
```




## Caveats with the data set

- The definitions of each column type are as follows:
  - **'name_in_reference'** is the specific spelling/capitalization used in the pubmed article
  - **'name_from_itcr_simplified'** is the input key provided to the database (case insensitive word matching)
  - **"usage"** is usage, **"allusion"** is a simple mention, like 'software xyz exists', 
  - **"deposition"** and **"creation"** indicate statements about the development or release of the software, like 'we put xyz on github'
  
- The software that generated the data considered the names as case-sensitive. So, for example, "Galaxy", "GALAXY" were considered distinct.

- Look at the following table for a specific article (ID: PMC4765054), 
  - **name_in_reference** (column 3) shows that issue mentioned above. 
  - However, **toolName** (column 7) is same for all 6 rows

- Therefore, to generate the frequency table for **how many articles** cited a specific software, we need only keep the rows that are distinct in toolName and article (PMC IDs). 

```{r issue1}
wdat %>% filter(article == "PMC4765054") %>% arrange(name_in_reference) %>% select(-software) %>% 
  DT::datatable(caption = htmltools::tags$caption("Case sensitivity and redundant rows", style = "color:red"))
```

- Keep in mind that, each article can mention two or more tools.

```{r issue2}
wdat %>% filter(article == "PMC4606501") %>% arrange(name_in_reference) %>% select(-software) %>% 
  DT::datatable(caption = htmltools::tags$caption("Example of one article mentioning multiple tools", style = "color:red"))
```

```{r johnsTable}
## As a measure of check,
# ## Re-create the first table sent by John (case-sensitive rows merged)
# wdat %>% 
#   group_by(toolName) %>% 
#   count(mentionType, wt = num_abs_mention) %>% 
#   ungroup() %>% 
#   pivot_wider(names_from = mentionType, values_from = n) 
# # %>% summarize(across(.cols = where(is.numeric), .fns = sum, na.rm = TRUE) ) %>% sum

# unique(wdat %>% select(article, toolName))
```

- There are **`r unique(wdat$article) %>% length()` unique articles** in the data set.

## Frequency distribution of article mentions for each tool
  
```{r frequencyTab}
## Overall frequency table (allusion or creation or deposition or usage)
freq_tab <- wdat %>% group_by(toolName) %>% count(article) %>% ungroup() %>% 
  count(toolName) 

freq_tab %>% arrange(desc(n)) %>% rename(`Total article mentions` = n) %>% 
  DT::datatable(caption = htmltools::tags$caption("Overall frequency table (allusion or creation or deposition or usage)", 
                                                  style = "color:blue"),
                options = list(columnDefs = list(list(className = 'dt-center', targets="_all"))))
```

## Frequency of different mode of mentions 

- Note that the last four columns (Allusion, Creation, Deposition, Usage) **Need Not sum** to the total article mentions as one or more of these different type of mentions may have occurred for a speicific tool in each article. 

```{r modesOfMentions}
freq_tab_full <- wdat %>% 
  group_by(toolName) %>% count(article, mentionType) %>% ungroup() %>% 
  group_by(mentionType) %>% count(toolName) %>% ungroup() %>% arrange(toolName) %>% 
  pivot_wider(names_from = mentionType, values_from = n) %>% 
  right_join(freq_tab, by = "toolName") %>% relocate(n, .after = toolName) %>% 
  arrange(desc(n)) %>% rename(`Total article mentions` = n) 

 freq_tab_full %>% 
  # %>% rowwise() %>% mutate(total = sum(Allusion+Creation+Deposition+Usage, na.rm = T))
  DT::datatable(caption = htmltools::tags$caption("Frequency of different mode of mentions", style = "color:blue"),
                options = list(columnDefs = list(list(className = 'dt-center', targets = "_all"))))
```

## REDEFINE total metntions using only (Allusion+Deposition+Usage)
```{r}
freq_tab_full_redefined <- wdat %>% 
  ## We filter out the Creations 
  filter(mentionType != "Creation") %>% 
  group_by(toolName) %>% count(article, mentionType) %>% ungroup() %>% 
  group_by(mentionType) %>% count(toolName) %>% ungroup() %>% arrange(toolName) %>% 
  pivot_wider(names_from = mentionType, values_from = n) %>% 
  right_join(freq_tab, by = "toolName") %>% relocate(n, .after = toolName) %>% 
  arrange(desc(n)) %>% rename(`Total article mentions` = n) 

## Use ^this freq_tab_full_redefined for the data merger
```


```{r mergerWithToolAnalysis}
## Join ToolAnalysis data with SoftMention data
## We consider QIIME 2 as QIIME for simplicity
## mwdat = merged working data
tdat_edited <- tdat %>% 
  mutate(## Extract all the acronyms inside the parentheses
         names_cut = str_extract_all(toolName, "\\([^()]+\\)") %>% substring(., 2, nchar(.)-1),
         ## remove the nonsensical names_cut for names that did not contain acronyms inside the parentheses
         names_cut = if_else(names_cut == "haracter(0", toolName, names_cut),
         ## Keep only the first word in each name (since this is most common in the Software data)
         names_cut = word(names_cut, 1),
         ## Individually fix those who had standalone issues or problems with operations above
         names_cut = case_when(names_cut == "AA" ~ "AmpliconArchitect",
                               names_cut == "AR" ~ "AmpliconReconstructor",
                               names_cut == "dcmqi:" ~ "dcmqi",
                               names_cut == "PDX" ~ "PDXFinder",
                               names_cut == "3D" ~ "3D Slicer",
                               names_cut == "Pathway" ~ "Pathway Commons",
                               names_cut == "UCSC" ~ "Xena",
                               TRUE ~ names_cut),
         ## remove any whitespaces before after names
         names_cut = str_trim(names_cut)) %>% 
  ## remove the original tool name variable and replace with newly generated names for merging
  select(-toolName) %>% rename(toolName = names_cut) %>% relocate(toolName) 

## merge with the smaller data set on the right (Update: Use the counts without creation)
mwdat <- tdat_edited %>% 
  right_join(freq_tab_full_redefined %>% rename(totalMentions = `Total article mentions`), by = "toolName")

## Some helpful recoding
mwdat %<>% mutate(citeHow = if_else(citeHow == "Yes", citeHow, "No"),
                 TimeSinceLastUpdate_bin = if_else(TimeSinceLastUpdate < 2, "<= 1 year", "2+ years"),
                 ## combine not working with limited: extensiveContact
                 extensiveContact = if_else(extensiveContact == "NotWorking", "None", extensiveContact)
                 )

#mwdat %>% count(extensiveContact)
```

## Analyses

### 1. Possible association between social media presence and log(totalMentions)

- Note that we are using log-scaled data since the data is count-type and there is considerable variation in it.

```{r socialMedia}
## Marginal dist
p1 <- mwdat %>% 
  ggplot(aes(x = socialMedia, fill = fct_infreq(socialMedia))) + 
  geom_bar(position = "dodge", width = .5, show.legend = F) + 
  geom_text(stat = 'count', 
            aes(label = paste0(after_stat(round(100*count/sum(count),1)), "%")), 
            position = position_dodge(width = .9), vjust = 2, col = "white") +
  labs(x = "Social Media Presence", y = "Frequency", subtitle = "Distribution of Social Media Presence") +
  scale_y_continuous(n.breaks = 7) + 
  scale_fill_brewer(palette = "Set1", direction = 1) +
  theme_light() 

p2 <- mwdat %>% 
  ggplot(aes(x = socialMedia, y = log(totalMentions))) +
  geom_boxplot(aes(fill = socialMedia),  width = 0.5, show.legend = F) +
  scale_y_continuous(n.breaks = 6) + 
  scale_fill_brewer(palette = "Set1", direction = 1) +
  labs(x = "Social Media Presence", y = "log( total # of mentions)", 
       subtitle = "Distribution of # of mentions (log scale)") +
  theme_light() 

p1 + p2
```

```{r}
## Allusions
p21 <- mwdat %>% filter(!is.na(Allusion)) %>% 
  ggplot(aes(x = socialMedia, y = log(Allusion))) +
  geom_boxplot(aes(fill = socialMedia),  width = 0.5, show.legend = F) +
  scale_y_continuous(n.breaks = 6) + 
  scale_fill_brewer(palette = "Set1", direction = 1) +
  labs(x = "Social Media Presence", y = "log(Allusion)", subtitle = "Distribution of # of Allusion (log scale)") +
  theme_light() + coord_flip()

## Deposition
p23 <- mwdat %>% filter(!is.na(Deposition)) %>% 
  ggplot(aes(x = socialMedia, y = log(Deposition))) +
  geom_boxplot(aes(fill = socialMedia),  width = 0.5, show.legend = F) +
  scale_y_continuous(n.breaks = 6) + 
  scale_fill_brewer(palette = "Set1", direction = 1) +
  labs(x = "Social Media Presence", y = "log(Deposition)", 
       subtitle = "Distribution of # of Deposition (log scale)") +
  theme_light() + coord_flip()

## Usage
p24 <- mwdat %>% filter(!is.na(Usage)) %>% 
  ggplot(aes(x = socialMedia, y = log(Usage))) +
  geom_boxplot(aes(fill = socialMedia),  width = 0.5, show.legend = F) +
  scale_y_continuous(n.breaks = 6) + 
  scale_fill_brewer(palette = "Set1", direction = 1) +
  labs(x = "Social Media Presence", y = "log(Usage)", subtitle = "Distribution of # of Usage (log scale)") +
  theme_light() + coord_flip()

(p21) / (p23 + p24)
```

- It appears the distribution of log-frequency (and hence the frequency) differs significantly. 
- However, mention is a function of time as well. So we need to check whether adjusting age of tool affects this crude association.

```{r warning=FALSE}
## I found out the release years for the tools: Cytoscape (2002), RDriver (2017)
regdat <- mwdat %>% 
  mutate(releaseYear = case_when(toolName == "rDriver" ~ 2017, 
                                 toolName == "Cytoscape" ~ 2002, 
                                 TRUE ~ releaseYear),
         ## Year 2022 is completed, hence those created in 2022 could be considered as Year 1 age
         toolAge = 2023 - releaseYear,
         ## Since this is only the beginning of 2023, we count up to the year 2022
         TimeSinceLastUpdate = 2022 - updateYear)

p3 <- regdat %>% 
  ggplot(aes(x = releaseYear, y = log(totalMentions), col = socialMedia)) + 
  geom_point(aes(shape = socialMedia)) + 
  geom_smooth(method = "lm", se = F) +
  labs(x = "Release Year", y = "log(total Mentions)", subtitle = "Scatterplot of log-totalMentions vs Release Year") +
  theme_light() 

p3
```


### An exploratory regression
```{r regression}
lm(log(totalMentions) ~ socialMedia, data = regdat) %>% broom::tidy() %>% 
  mutate(across(.cols = where(is.numeric), .fns = ~round(.x, digits = 4))) %>% 
  DT::datatable()

lm(log(totalMentions) ~ socialMedia + toolAge, data = regdat) %>% broom::tidy() %>% 
  mutate(across(.cols = where(is.numeric), .fns = ~round(.x, digits = 5))) %>% 
  DT::datatable()
```

- Adjusting for Tool age (a translation of release Year) shows that social media presence is not significant. 
- The variability is too high to conclude anything statistically. 


## Stratification by: How to cite instruction (with BioConductor)

```{r Howtocite}
## Marginal dist
p4 <- mwdat %>% 
  ggplot(aes(x = citeHow, fill = fct_infreq(citeHow))) + 
  geom_bar(position = "dodge", width = .5, show.legend = F) + 
  geom_text(stat = 'count', 
            aes(label = paste0(after_stat(round(100*count/sum(count),1)), "%")), 
            position = position_dodge(width = .9), vjust = 2, col = "white") +
  labs(x = "Instruction for citation", y = "Frequency", subtitle = "Distribution of Presence of Citation instructions") +
  scale_y_continuous(n.breaks = 7) + 
  scale_fill_brewer(palette = "Set1", direction = -1) +
  theme_light() 

p5 <- mwdat %>% 
  ggplot(aes(x = citeHow, y = log(totalMentions))) +
  geom_boxplot(aes(fill = citeHow),  width = 0.5, show.legend = F) +
  scale_y_continuous(n.breaks = 6) + 
  scale_fill_brewer(palette = "Set1", direction = 1) +
  labs(x = "Instruction for citation", y = "log( total # of mentions)", 
       subtitle = "Distribution of # of mentions (log scale)") +
  theme_light() 

p4 + p5
```


```{r}
regdat %>% 
  ggplot(aes(x = citeHow, y = log(totalMentions))) +
  geom_boxplot(aes(fill = socialMedia),  width = 0.5, ) +
  scale_y_continuous(n.breaks = 6) + 
  #facet_wrap(~`class/type`) +
  scale_fill_brewer(palette = "Set1", direction = 1) +
  labs(x = "Instructions on how to cite?", y = "log( total # of mentions)", 
       subtitle = "Distribution of # of mentions (log scale)") +
  theme_light() +
  theme(strip.background =element_rect(fill="steelblue3"),
        strip.text = element_text(colour = 'white'), 
        panel.grid = element_blank())

lm(log(totalMentions) ~ socialMedia + citeHow + socialMedia:citeHow, data = regdat) %>% broom::tidy() %>% 
  mutate(across(.cols = where(is.numeric), .fns = ~round(.x, digits = 5))) %>% 
  DT::datatable()
```

- Note: It seems that Tool age is a very strong predictor of the log(mentions), and makes the other variables insignificant.

## (coming) Stratification by: How to cite instruction (WITHOUT BioConductor and other-R packages)

```{r Howtocite_WITHOUT}
## Marginal dist
p41 <- mwdat %>% filter(`class/type` != c("Bioconductor R packages", "other R packages")) %>% 
  ggplot(aes(x = citeHow, fill = fct_infreq(citeHow))) + 
  geom_bar(position = "dodge", width = .5, show.legend = F) + 
  geom_text(stat = 'count', 
            aes(label = paste0(after_stat(round(100*count/sum(count),1)), "%")), 
            position = position_dodge(width = .9), vjust = 1.2, col = "white") +
  labs(x = "Instruction for citation", y = "Frequency", subtitle = "Distribution of Presence of Citation instructions") +
  scale_y_continuous(n.breaks = 7) + 
  scale_fill_brewer(palette = "Set1", direction = -1) +
  theme_light() 

p51 <- mwdat %>% filter(`class/type` != c("Bioconductor R packages", "other R packages")) %>% 
  ggplot(aes(x = citeHow, y = log(totalMentions))) +
  geom_boxplot(aes(fill = citeHow),  width = 0.5, show.legend = F) +
  scale_y_continuous(n.breaks = 6) + 
  scale_fill_brewer(palette = "Set1", direction = 1) +
  labs(x = "Instruction for citation", y = "log( total # of mentions)", 
       subtitle = "Distribution of # of mentions (log scale)") +
  theme_light() 


p61 <- regdat %>% filter(`class/type` != c("Bioconductor R packages", "other R packages")) %>% 
  ggplot(aes(x = citeHow, y = log(totalMentions))) +
  geom_boxplot(aes(fill = socialMedia),  width = 0.5, ) +
  scale_y_continuous(n.breaks = 6) + 
  #facet_wrap(~`class/type`) +
  scale_fill_brewer(palette = "Set1", direction = 1) +
  labs(x = "Instructions on how to cite?", y = "log( total # of mentions)", 
       subtitle = "Distribution of # of mentions (log scale)") +
  theme_light() +
  theme(strip.background =element_rect(fill="steelblue3"),
        strip.text = element_text(colour = 'white'), 
        panel.grid = element_blank())

(p41 + p51) /
  p61
```

## Stratification by: extensiveContact

```{r}
regdat %>% mutate(extensiveContact = if_else(extensiveContact == "None", extensiveContact, "Yes")) %>% 
  ggplot(aes(x = extensiveContact, y = log(totalMentions))) +
  geom_boxplot(aes(fill = extensiveContact), width = 0.4, show.legend = F) +
  scale_y_continuous(n.breaks = 6) + 
  #facet_wrap(~`class/type`) +
  scale_fill_brewer(palette = "Set1", direction = 1) +
  labs(x = "Extensive Contact method availability", y = "log( total # of mentions)", 
       subtitle = "Distribution of # of mentions (log scale)") +
  theme_light() +
  theme(strip.background =element_rect(fill="steelblue3"),
        strip.text = element_text(colour = 'white'), 
        panel.grid = element_blank()) + coord_flip()

regdat %>% mutate(extensiveContact = if_else(extensiveContact == "None", extensiveContact, "Yes")) %>% 
lm(log(totalMentions) ~ extensiveContact + toolAge, data = .) %>% broom::tidy() %>% 
  mutate(across(.cols = where(is.numeric), .fns = ~round(.x, digits = 5))) %>% 
  DT::datatable()
```

- Having Extensive contact method available is significant EVEN AFTER Ajdusting for Tool age.

## (NEW) Additional Plot: Diversity of the type of tools
```{r}
regdat %>% select(toolName, toolType = `class/type`) %>% 
  ggplot(aes(y = fct_infreq(toolType), fill = toolType)) +
  geom_bar(width = .5, show.legend = FALSE) + 
  labs(x = "Frequency", y = "Type of tool") +
  scale_x_continuous(n.breaks = 7) + 
  #scale_fill_brewer(palette = "Set3", direction = 1) +
  theme_light() +
  theme(panel.grid = element_blank() ) 
```

## (coming) Overlap of tools between survey data and Tool Analysis

```{r}
tdat_edited %>% mutate(ToolAn_Data = toolName) %>% select(toolName, ToolAn_Data) %>% 
  left_join(freq_tab_full_redefined %>% mutate(SImpact_Data = toolName), by = "toolName") %>% 
  relocate(ToolAn_Data, SImpact_Data) %>% select(-toolName) %>% 
  arrange(SImpact_Data) %>% 
  filter(!is.na(SImpact_Data)) %>% 
  DT::datatable(caption = "These tools had information in the Software Impact Analysis dataset")

tdat_edited %>% mutate(ToolAn_Data = toolName) %>% select(toolName, ToolAn_Data) %>% 
  left_join(freq_tab_full_redefined %>% mutate(SImpact_Data = toolName), by = "toolName") %>% 
  relocate(ToolAn_Data, SImpact_Data) %>% select(-toolName) %>% 
  arrange(SImpact_Data) %>% 
  filter(is.na(SImpact_Data)) %>% 
  mutate(ToolAn_Data = case_when(ToolAn_Data == "Gene" ~ "Gene Pattern Notebook",
                                 TRUE ~ ToolAn_Data)) %>% 
  DT::datatable(caption = "These tools had no information in the Software Impact Analysis dataset")
```


## Documentation Depth (manual data)

- The information on all the tools regarding the extensiveness of the documentation was manually collected.

```{r}
#regdat %>% pull(toolName)

regdat %<>% 
  mutate(DocuDepth = case_when(
    toolName == "Galaxy" ~ "Extensive",
    toolName == "3D Slicer" ~ "Extensive",
    toolName == "CRAVAT" ~ "Extensive",
    toolName == "geWorkbench" ~ "Extensive",
    toolName == "QIIME" ~ "Extensive",
    toolName == "cBioPortal" ~ "Extensive",
    toolName == "Xena" ~ "Webpage/PDF/ReadMe",
    toolName == "DepMap" ~ "Extensive",
    toolName == "TCPA" ~ "Extensive",
    toolName == "OncoPPi" ~ "Webpage/PDF/ReadMe",
    toolName == "JBrowse" ~ "Extensive",
    toolName == "PhenStat" ~ "Extensive",
    toolName == "Bioconductor" ~ "Extensive",
    toolName == "ELMER" ~ "Extensive",
    toolName == "DIGGIT" ~ "Webpage/PDF/ReadMe",
    toolName == "MethylMix" ~ "Webpage/PDF/ReadMe",
    toolName == "VIPER" ~ "Extensive",
    toolName == "AMARETTO" ~ "Extensive",
    toolName == "CORE" ~ "Extensive",
    toolName == "InGRiD" ~ "Webpage/PDF/ReadMe",
    toolName == "DANA" ~ "Webpage/PDF/ReadMe",
    toolName == "AmpliconReconstructor" ~ "Webpage/PDF/ReadMe",
    toolName == "AmpliconArchitect" ~ "Webpage/PDF/ReadMe",
    toolName == "HistoQC" ~ "Webpage/PDF/ReadMe",
    toolName == "GSEA" ~ "Extensive",
    toolName == "Cytoscape" ~ "Extensive",
    toolName == "Trinity" ~ "Webpage/PDF/ReadMe",
    toolName == "TIES" ~ "Webpage/PDF/ReadMe", ## NONE
    toolName == "OncoMX" ~ "Extensive",
    toolName == "MSigDB" ~ "Webpage/PDF/ReadMe",
    toolName == "Pathway Commons" ~ "Extensive",
    toolName == "PDXFinder" ~ "Webpage/PDF/ReadMe", ## NONE
    toolName == "rDriver" ~ "Webpage/PDF/ReadMe",
    toolName == "dcmqi" ~ "Extensive",
    toolName == "Lancet" ~ "Webpage/PDF/ReadMe",
    toolName == "D2Refine" ~ "Webpage/PDF/ReadMe"),
    DocuDepth = fct_relevel(DocuDepth, "Webpage/PDF/ReadMe"))
```


```{r}
## Marginal dist
p_doc1 <- regdat %>%  
  ggplot(aes(x = DocuDepth, fill = fct_infreq(DocuDepth))) + 
  geom_bar(position = "dodge", width = .5, show.legend = F) + 
  geom_text(stat = 'count', 
            aes(label = paste0(after_stat(round(100*count/sum(count),1)), "%")), 
            position = position_dodge(width = .9), vjust = 2, col = "white") +
  labs(x = "Documentation", y = "Frequency", subtitle = "Distribution of Presence of Documentation depth") +
  scale_y_continuous(n.breaks = 7) + 
  scale_fill_brewer(palette = "Set1", direction = -1) +
  theme_light() 

p_doc2 <- regdat %>%
  ggplot(aes(x = DocuDepth, y = log(totalMentions), fill = DocuDepth)) +
  geom_boxplot(width = 0.5, show.legend = T) +
  scale_y_continuous(n.breaks = 6) + 
  scale_fill_brewer(palette = "Set1", direction = 1) +
  labs(x = "Documentation depth", y = "log( total # of mentions)", 
       subtitle = "Distribution of # of mentions (log scale)") +
  theme_light() 

p_doc1 + p_doc2
```

```{r}
lm(log(totalMentions) ~ DocuDepth, data = regdat) %>% broom::tidy() %>% 
  mutate(across(.cols = where(is.numeric), .fns = ~round(.x, digits = 5))) %>% 
  DT::datatable()

regdat %>% mutate(extensiveContact = if_else(extensiveContact == "None", extensiveContact, "Yes")) %>% 
lm(log(totalMentions) ~ extensiveContact + DocuDepth + toolAge, data = .) %>% broom::tidy() %>% 
  mutate(across(.cols = where(is.numeric), .fns = ~round(.x, digits = 5))) %>% 
  DT::datatable()

lm(log(totalMentions) ~ DocuDepth + socialMedia + extensiveContact + toolAge, data = regdat) %>% broom::tidy() %>% 
  mutate(across(.cols = where(is.numeric), .fns = ~round(.x, digits = 5))) %>% 
  DT::datatable()

#names(regdat)
lm(log(Usage) ~ toolAge +simple_health_metrics*socialMedia, data = filter(regdat, toolAge >=3) ) %>% broom::tidy() %>% 
  mutate(across(.cols = where(is.numeric), .fns = ~round(.x, digits = 5))) %>% 
  DT::datatable()

```

```{r CarriePlots, fig.height=8, fig.width=8}
# Social Media

regdat %<>% mutate(
  socialMedia = factor(socialMedia, levels = c("Yes", "No")),
  
)
d1 <- regdat %>% 
  ggplot(aes(x = releaseYear, y = log(Usage), col = socialMedia)) + 
  geom_point(aes(shape = socialMedia)) + 
  geom_smooth(method = "lm", se = T, alpha = 0.2) +
  labs(x = "Release Year", y = "log(total Usage)", title = "Social Media Presence") +
  theme_light() +coord_equal() + theme(legend.position = "bottom")

## DocuDepth
d2 <- regdat %>%
  ggplot(aes(x = releaseYear, y = log(Usage), col = DocuDepth)) + 
  geom_point(aes(shape = DocuDepth)) + 
  geom_smooth(method = "lm", se = T, alpha = 0.2) +
  labs(x = "Release Year", y = "log(total Usage)", title = "Documentation Depth") +
  theme_light() +coord_equal() + theme(legend.position = "bottom")

# extensiveContact
d3 <- regdat %>% 
  ggplot(aes(x = releaseYear, y = log(Usage), col = extensiveContact)) + 
  geom_point(aes(shape = extensiveContact)) + 
  geom_smooth(method = "lm", se = T, alpha = 0.2) +
  labs(x = "Release Year", y = "log(total Usage)", title = "Contact Depth") +
  theme_light() +coord_equal() + theme(legend.position = "bottom")

# CiteHow
d4 <- regdat %>% 
  ggplot(aes(x = releaseYear, y = log(Usage), col = citeHow)) + 
  geom_point(aes(shape = citeHow)) + 
  geom_smooth(method = "lm", se = T, alpha = 0.2) +
  labs(x = "Release Year", y = "log(total Usage)", subtitle = "Information for citing") +
  theme_light() +coord_equal() + theme(legend.position = "bottom")

# health metrics
d5 <- regdat %>% drop_na(simple_health_metrics) %>%
  ggplot(aes(x = releaseYear, y = log(Usage), col = simple_health_metrics)) + 
  geom_point(aes(shape = simple_health_metrics)) + 
  geom_smooth(method = "lm", se = T, alpha = 0.2) +
  labs(x = "Release Year", y = "log(total Usage)", title = "Presence of Software Health Metrics", subtitle = ("test coverage and build status badges")) +
  theme_light() +coord_equal() + theme(legend.position = "bottom")

(d1 + d2) / (d3 + d5)
```

```{r CarriePlots2, fig.height=8, fig.width=8}
# Social Media
d11 <- regdat %>% 
  ggplot(aes(x = releaseYear, y = log(totalMentions), col = socialMedia)) + 
  geom_point(aes(shape = socialMedia)) + 
  geom_smooth(method = "lm", se = T, alpha = 0.2) +
  labs(x = "Release Year", y = "log(Ttotal Mentions)", title = "Social Media Presence") +
  theme_light() +coord_equal() + theme(legend.position = "bottom")

## DocuDepth
d22 <- regdat %>%
  ggplot(aes(x = releaseYear, y = log(totalMentions), col = DocuDepth)) + 
  geom_point(aes(shape = DocuDepth)) + 
  geom_smooth(method = "lm", se = T, alpha = 0.2) +
  labs(x = "Release Year", y = "log(Total Mentions)", title = "Documentation Depth") +
  theme_light() +coord_equal() + theme(legend.position = "bottom")

# extensiveContact
d33 <- regdat %>% 
  ggplot(aes(x = releaseYear, y = log(totalMentions), col = extensiveContact)) + 
  geom_point(aes(shape = extensiveContact)) + 
  geom_smooth(method = "lm", se = T, alpha = 0.2) +
  labs(x = "Release Year", y = "log(Total Mentions)", title =  "Contact Depth") +
  theme_light() +coord_equal() + theme(legend.position = "bottom")

# CiteHow
d44 <- regdat %>% 
  ggplot(aes(x = releaseYear, y = log(totalMentions), col = citeHow)) + 
  geom_point(aes(shape = citeHow)) + 
  geom_smooth(method = "lm", se = T, alpha = 0.2) +
  labs(x = "Release Year", y = "log(Total Mentions)", subtitle = "Information on how to cite") +
  theme_light() +coord_equal() + theme(legend.position = "bottom")

d55<- regdat %>% drop_na(simple_health_metrics) %>%
  ggplot(aes(x = releaseYear, y = log(totalMentions), col = simple_health_metrics)) + 
  geom_point(aes(shape = simple_health_metrics)) + 
  geom_smooth(method = "lm", se = T, alpha = 0.2) +
  labs(x = "Release Year", y = "log(Total Mentions)", title = "Presence of Software Health Metrics", subtitle = ("test coverage and build status badges")) +
  theme_light() +coord_equal() + theme(legend.position = "bottom")

(d11 + d22) / (d33 + d55)
```

```{r}
regdat  %>% 
  ggplot(aes(x = toolAge)) +
  geom_histogram(aes(y = after_stat(density)), fill = "white", col = "darkgray", binwidth = 2) +
  geom_density(col = "steelblue2") +
  scale_x_continuous(breaks = seq(0, 22, by = 2)) +
  labs(x = "Tool Age", y = "Density", subtitle = "Tool Age distribution") +
  theme_light() + 
  theme(
    panel.grid = element_blank()
  )
```


## Freq table for type of Social Media

```{r}
regdat %>% count(Twitter.bi, Youtube.bi, OtherMedia.bi)
```






