---
title: "Tool Analysis"
subtitle: "Initial questions"
author: "Awan"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_download: true 
    code_folding: 'hide'
    theme: cosmo
    highlight: pygments
---

<style type="text/css">
body .main-container {
  max-width: 1200px !important;
  width: 1200px !important;
}
body {
  max-width: 1200px !important;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,  fig.align='center')
## Required packages
library(tidyverse)
library(magrittr)
library(googlesheets4)
library(DT)
library(scales)
# install.packages("devtools")
library(patchwork) ## plot side by side
```


## Part I: Data loading and check

```{r loadData, eval=FALSE, echo=TRUE}
## Load data file and fix the Year column entries which appear as lists
tdat <- read_sheet(ss = "https://docs.google.com/spreadsheets/d/1-GLClAEl0aydwQEX-EeHiVufIhZCvervXc35fBAQWIc/edit?usp=sharing") %>% 
  mutate(`year of 1st release` = `year of 1st release` %>% as.character,
         `year of the most recent version/update` = `year of the most recent version/update` %>% as.character)

## For first time use, log in and give access to Tidyverse API Packages to [See, edit, create and delete all your Google Sheets spreadsheets]. 
## When running after first authentication, run "1" to make the userID selection in the console
write_csv(x = tdat, file = here::here("tdat.csv"))
```

### Loading data
```{r}
wdat <- read_csv(file = "tdat.csv") 
```

## Dichotomize the variables for analyses
```{r}
wdat %<>% 
  #rename("ownDOI" = "Have a DOI for the software itself?") %>% 
  mutate(ownDOI = if_else(!is.na(`Have a DOI for the software itself?`), true = "Yes", "No"),
         ## Social Media
         Twitter.bi = if_else(!is.na(`Twitter`), true = "Yes", "No"),
         Youtube.bi = if_else(!is.na(`Youtube`), true = "Yes", "No"),
         OtherMedia.bi = if_else(!is.na(`Other media`), true = "Yes", "No"),
         ## Any social Media? 
         socialMedia = if_else(Twitter.bi == "Yes" | Youtube.bi == "Yes" | OtherMedia.bi == "Yes", true = "Yes", "No"),
         ## Instructions of How to Cite? 
         citeHow = case_when(str_detect(`Instructions how to cite`, 'Yes') ~ "Yes",
                             str_detect(`Instructions how to cite`, 'unclear') ~ "Unclear",
                             TRUE ~ `Instructions how to cite`),
         ###  Those without Yes/No/Unclear citation instructions to be included in "Yes"
         citeHow = case_when(str_detect(citeHow, 'Yes|No|Unclear') ~ citeHow,
                             TRUE ~ "Yes"))
```


### Fixing Release and Update date entries + Calculate ToolAge, Elapsed time since last update
```{r}
## Release Issue 1: Gene Pattern Notebook - Which one to consider as start/release year. 
## Gene Pattern release 2006, but Notebook release 2017
## I used 2006 as suggested. 

## Update issue 1: geWorkbench - most recent version 2016 (web page updated in 2022). Which one to consider? 
##  I used 2022. (as suggested by Carrie: "I think update is the wrong word here - maybe maintenance date (I know that is what Andrey put, but update kinda implies software update) - but yes, 2022 seems reasonable - I think they are keeping up with maintaining it")

## Update issue 2:  Text Information Extraction System (TIES) - this software and forum are no longer maintained, as of 8/15/2019. 
## I used 2019 (Carrie:  would say 2019 is the most recent date)

wdat %<>% 
  #select(`tool name`, `year of 1st release`, `year of the most recent version/update`) %>% 
  mutate(## Release Year 
         releaseYear = if_else(`tool name` == "Gene Pattern Notebook", true = "2006", `year of 1st release`),
         releaseYear = if_else(str_detect(releaseYear, pattern = "2"), as.numeric(releaseYear), NA_real_),
         ## I found out the release years for the tools: Cytoscape (2002), RDriver (2017)
         releaseYear = case_when(`tool name` == "rDriver" ~ 2017, 
                                 `tool name` == "Cytoscape" ~ 2002, 
                                 TRUE ~ releaseYear),
         ## Latest Update Year
         updateYear = case_when(`tool name` == "geWorkbench" ~ 2022,
                                `tool name` == "Text Information Extraction System (TIES)" ~ 2019,
                                `year of the most recent version/update` == "unknown" ~ NA_real_,
                                TRUE ~ as.numeric(`year of the most recent version/update`)),
         ## Year 2022 is completed, hence those created in 2022 could be considered as Year 1 age
         toolAge = 2023 - releaseYear,
         ## Since this is only the beginning of 2023, we count up to the year 2022
         TimeSinceLastUpdate = 2022 - updateYear)
```


## Recoding other interesting variables
1. extensive contact/feedback 
2. software health metrics (unit testing, automation, code coverage stats)

```{r}
wdat %<>% 
  #select(`tool name`, `extensive contact/feedback`, `software health metrics (unit testing, automation, code coverage stats)`) %>% 
  mutate(## extensive contact/feedback 
         extensiveContact = case_when(str_detect(`extensive contact/feedback`, "not working|no posts") ~ "NotWorking",
                                      `extensive contact/feedback` == "No" ~ "None", 
                                      TRUE ~ "Yes"),
         ## software health metrics
         softHealthMetrics = str_sub(`software health metrics (unit testing, automation, code coverage stats)`, 1, 14), ## keep first 14 chars
         softHealthMetrics = str_replace(`softHealthMetrics`, "statistic|stati|st", "stats"),  ##replace with meaninful string "stats"
         softHealthMetrics = if_else(softHealthMetrics == "Limited by num", "Installation count", softHealthMetrics), ## Better recode
         softHealthMetrics = if_else(str_detect(softHealthMetrics, "Download"), "Download stats", softHealthMetrics), ## Better recode
         softHealthMetrics = str_trim(softHealthMetrics, "both") ## Remove white space that may exist
         ) 

# Saving edited dataset
write_csv(x = wdat, file = here::here("tdat_edited.csv"))
```

# Graphical Analyses

## 1. Proportion of softwares with DOI
```{r, fig.align='center'}
wdat %>% 
  ggplot(aes(x = ownDOI, fill = ownDOI)) +
  geom_bar( width = 0.5) + 
  geom_text(stat = 'count', aes(label = paste0(after_stat(round(100*count/sum(count),1)), "%")), vjust = 1.5) +
  labs(x = "softwares with DOI", y = "Frequency", caption = "Distribution of softwares with own DOI") +
  scale_y_continuous(n.breaks = 10) + 
  scale_fill_brewer(palette = "Set2", direction = -1) +
  theme_light() 

wdat %>% 
  ggplot(aes(x = ownDOI, fill = ownDOI)) +
  geom_bar( width = 0.5) + 
  geom_text(stat = 'count', aes(label = after_stat(count), vjust = 1.5)) +
  facet_wrap(~`class/type`) +
  labs(x = "softwares with DOI", y = "Frequency", caption = "Distribution of softwares with own DOI") +
  scale_y_continuous(n.breaks = 10) + 
  scale_fill_brewer(palette = "Set2", direction = -1) +
  theme_light() +
  theme(strip.background =element_rect(fill="steelblue3"),
        strip.text = element_text(colour = 'white'), 
        panel.grid = element_blank())
```


## 2. Any kind of social media presence
```{r, fig.align='center'}
## Marginal dist
wdat %>% 
  ggplot(aes(x = socialMedia, fill = fct_infreq(socialMedia))) + 
  geom_bar(position = "dodge", width = .6, show.legend = F) + 
  geom_text(stat = 'count', 
            aes(label = paste0(after_stat(round(100*count/sum(count),1)), "%")), position = position_dodge(width = .9), vjust = 1.5) +
  labs(x = "Social Media Presence", y = "Frequency", fill = "Social Media Presence", caption = "Distribution of Social Media Presence") +
  scale_y_continuous(n.breaks = 7) + 
  scale_fill_brewer(palette = "Set2", direction = -1) +
  theme_light() 

## Stratified by
wdat %>% 
  ggplot(aes(x = socialMedia, fill = fct_infreq(socialMedia))) + 
  geom_bar(position = "dodge", width = .6, show.legend = F) + 
  geom_text(stat = 'count', 
            aes(label = paste0(after_stat(count))), position = position_dodge(width = .9), vjust = 1.5) +
  facet_wrap(~`class/type`) +
  labs(x = "Social Media Presence", y = "Frequency", fill = "Social Media Presence", caption = "Distribution of Social Media Presence") +
  scale_y_continuous(n.breaks = 7) + 
  scale_fill_brewer(palette = "Set2", direction = -1) +
  theme_light() + 
  theme(strip.background =element_rect(fill="steelblue3"),
        strip.text = element_text(colour = 'white'), 
        panel.grid = element_blank())


## Distribution of type of social media
wdat %>% 
  select(socialMedia, Twitter.bi, Youtube.bi, OtherMedia.bi, `class/type`) %>% 
  pivot_longer(cols = -c(socialMedia, `class/type`), names_to = "socMed", values_to = "vals") %>% 
  filter(vals == "Yes") %>% 
  ggplot(aes(x = fct_infreq(socMed), fill = fct_infreq(socMed))) + 
  geom_bar(position = "dodge", width = .6, show.legend = F) + 
  #geom_text(stat = 'count', aes(label = paste0(after_stat(round(100*count/sum(count),2)), "%")), position = position_dodge(width = .9), vjust = 1.5) +
  facet_wrap(~`class/type`) +
  labs(x = "Social Media Outlets", y = "Frequency", fill = "Media Outlets", caption = "Distribution of Social Media Outlets (bi = Binary)") +
  scale_y_continuous(n.breaks = 7) + 
  scale_fill_brewer(palette = "Set2", direction = 1) +
  theme_light() +
  theme(strip.background =element_rect(fill="steelblue3"),
        strip.text = element_text(colour = 'white'),
        panel.grid = element_blank())
```


## 3. Proportion of How to cite 
```{r, fig.align='center'}
wdat %>% 
  ggplot(aes(x = fct_infreq(citeHow), fill = citeHow)) +
  geom_bar(position = "dodge", width = .5, show.legend = F) + 
  geom_text(stat = 'count', 
            aes(label = paste0(after_stat(round(100*count/sum(count),1)), "%")), position = position_dodge(width = .9), vjust = 1.5) +
  labs(x = "Instruction availability on How to Cite", y = "Frequency", caption = "Distribution of Citation Instructions") +
  scale_y_continuous(n.breaks = 12) + 
  scale_fill_brewer(palette = "Set2", direction = -1) +
  theme_light() 

## Stratified 
wdat %>% 
  ggplot(aes(x = fct_infreq(citeHow), fill = citeHow)) +
  geom_bar(position = "dodge", width = .5, show.legend = F) + 
  geom_text(stat = 'count', 
            aes(label = paste0(after_stat(count))), position = position_dodge(width = .9), vjust = 1.2) +
  facet_wrap(~`class/type`) +
  labs(x = "Instruction availability on How to Cite", y = "Frequency", caption = "Distribution of Citation Instructions") +
  #scale_y_continuous(n.breaks = 12) + 
  scale_fill_brewer(palette = "Set2", direction = -1) +
  theme_light() +
  theme(strip.background =element_rect(fill="steelblue3"),
        strip.text = element_text(colour = 'white'), 
        panel.grid = element_blank())
```


## 4. Age of the tool
```{r, fig.align='center'}
## Tool Age distribution
wdat %>% select(releaseYear) %>% mutate(releaseYear = as.factor(releaseYear)) %>% 
  ggplot(aes(x = releaseYear, fill = releaseYear)) +
  geom_bar(position = "dodge", width = .5, show.legend = F) + 
  labs(x = "Release Year", y = "Frequency", caption = "Distribution of Tool releases across the years") +
  scale_y_continuous(n.breaks = 7) + 
  #scale_fill_brewer(palette = "Set3", direction = 1) +
  theme_light() 

wdat %>%
  ggplot(aes(x = toolAge)) +
  geom_histogram(aes(y = after_stat(density)), fill = "white", col = "darkgray", binwidth = 2) +
  geom_density(col = "steelblue2") +
  scale_x_continuous(breaks = seq(0, 22, by = 2)) +
  labs(x = "Tool Age", y = "Density", subtitle = "Tool Age distribution") +
  theme_light() + 
  theme(
    panel.grid = element_blank()
  )
```
## Time elapsed since last update
```{r, fig.align='center'}
p1 <- wdat %>%
  ggplot(aes(x = TimeSinceLastUpdate)) +
  geom_histogram(aes(y = after_stat(density)), fill = "white", col = "darkgray", binwidth = 1) +
  geom_density(col = "steelblue2") +
  scale_x_continuous(breaks = seq(0, 6, by = 1)) +
  labs(x = "Elapsed time (years)", y = "Density", subtitle = "Time elaspsed since last update (those updated in 2022 is considered 0 time elapsed)") +
  theme_light() + 
  theme(panel.grid = element_blank() )

 ## Pattern in update tendencies
p2 <- wdat %>%
  ggplot(aes(x = toolAge, y = TimeSinceLastUpdate)) +
  geom_point(col = "navyblue", shape = 2) + 
  geom_smooth(se = F, col = "steelblue2") + 
  labs(y = "Elapsed time since last update (years)", x = "Tool Age (years)", caption = "No apparent pattern presents itself here") +
  scale_x_continuous(n.breaks = 12) +
  theme_light() + 
  theme(panel.grid = element_blank() )

p1 / p2

## Time elapsed since last update - One year vs more
wdat %>% 
  mutate(TimeSinceLastUpdate_bin = if_else(TimeSinceLastUpdate < 2, "<= 1 year", "2+ years")) %>% 
  ggplot(aes(x = TimeSinceLastUpdate_bin, fill = TimeSinceLastUpdate_bin)) +
  geom_bar(position = "dodge", width = .5, show.legend = F) + 
  labs(x = "Elapsed time category", y = "Frequency", caption = "Time elaspsed since last update (those updated in 2022 is considered 0 time elapsed)") +
  scale_y_continuous(n.breaks = 7) + 
  #scale_fill_brewer(palette = "Set1", direction = 1) +
  theme_light() 

## Time elapsed since last update - One year vs more || Stratified
wdat %>% 
  mutate(TimeSinceLastUpdate_bin = if_else(TimeSinceLastUpdate < 2, "<= 1 year", "2+ years")) %>% 
  #filter(!is.na(TimeSinceLastUpdate_bin) ) +
  ggplot(aes(x = TimeSinceLastUpdate_bin, fill = TimeSinceLastUpdate_bin)) +
  geom_bar(position = "dodge", width = .5, show.legend = F) + 
  labs(x = "Elapsed time category", y = "Frequency", caption = "Time elaspsed since last update (those updated in 2022 is considered 0 time elapsed)") +
  facet_wrap(~`class/type`) +
  scale_y_continuous(n.breaks = 7) + 
  #scale_fill_brewer(palette = "Set1", direction = 1) +
  theme_light() +
  theme(strip.background =element_rect(fill="steelblue3"),
        strip.text = element_text(colour = 'white'), 
        panel.grid = element_blank())
```

## 5. Software Health metrics

```{r, fig.align='center'}
wdat %>% select(`tool name`, softHealthMetrics, `class/type`) %>% 
  ggplot(aes(y = fct_infreq(softHealthMetrics), fill = softHealthMetrics)) +
  geom_bar(position = "dodge", width = .5, show.legend = F) + 
  geom_text(stat = 'count', 
            aes(label = paste0(after_stat(count))), position = position_dodge(width = .9), hjust = 1.5) +
  labs(x = "Software Health metrics", y = "Frequency", caption = "Distribution of Software Health metrics") +
  scale_x_continuous(n.breaks = 12) + 
  scale_fill_brewer(palette = "Set2", direction = -1) +
  theme_light()

wdat %>% select(`tool name`, softHealthMetrics, `class/type`) %>% 
  ggplot(aes(y = fct_infreq(softHealthMetrics), fill = softHealthMetrics)) +
  geom_bar(position = "dodge", width = .5, show.legend = F) + 
  geom_text(stat = 'count', 
            aes(label = paste0(after_stat(count))), position = position_dodge(width = .9), hjust = 1.5) +
  facet_wrap(~`class/type`) +
  labs(x = "Software Health metrics", y = "Frequency", caption = "Distribution of Software Health metrics") +
  #scale_x_continuous(n.breaks = 12) + 
  scale_fill_brewer(palette = "Set2", direction = -1) +
  theme_light() + 
  theme(strip.background =element_rect(fill="steelblue3"),
        strip.text = element_text(colour = 'white'))
```

## 6. Extensive contact/feedback
```{r}
wdat %>%
  ggplot(aes(y = fct_infreq(extensiveContact), fill = extensiveContact)) +
  geom_bar(position = "dodge", width = .5, show.legend = F) + 
  geom_text(stat = 'count', 
            aes(label = paste0(after_stat(round(100*count/sum(count),1)), "%")), position = position_dodge(width = .9), hjust = 1.2) +
  labs(x = "Extensive contact/feedback", y = "Frequency", caption = "Distribution of Extensive contact/feedback") +
  scale_x_continuous(n.breaks = 13) + 
  scale_fill_brewer(palette = "Set1", direction = 1) +
  theme_light() 

## stratified by tool type
wdat %>%
  ggplot(aes(y = fct_infreq(extensiveContact), fill = extensiveContact)) +
  geom_bar(position = "dodge", width = .5, show.legend = F) + 
  geom_text(stat = 'count', 
            aes(label = paste0(after_stat(count))), position = position_dodge(width = .9), hjust = 1.5) +
  facet_wrap(~`class/type`) + 
  labs(x = "Extensive contact/feedback", y = "Frequency", caption = "Distribution of Extensive contact/feedback") +
  #scale_x_continuous(n.breaks = 13) + 
  scale_fill_brewer(palette = "Set1", direction = 1) +
  theme_light() +
  theme(strip.background =element_rect(fill="steelblue3"),
        strip.text = element_text(colour = 'white'))
```

## 7. Check if distributions differ across different time categories

```{r}
wdat %>% filter(!is.na(releaseYear)) %>% 
  mutate(relYear.cut = cut_interval(releaseYear, 4)) %>% 
  ggplot(aes(fill = fct_infreq(extensiveContact), x = relYear.cut)) +
  geom_bar(position = "dodge", width = .5) + 
  # geom_text(stat = 'count', 
  #           aes(label = paste0(after_stat(round(100*count/sum(count),1)), "%")), position = position_dodge(width = .9), hjust = -0.2) +
  labs(x = "Release Year", y = "Frequency", fill = "Extensive contact/feedback") +
  scale_y_continuous(n.breaks = 13) + 
  scale_fill_brewer(palette = "Set2", direction = 1) +
  theme_light() 

# Stratified
wdat %>% filter(!is.na(releaseYear)) %>% 
  mutate(relYear.cut = cut_interval(releaseYear, 4)) %>% 
  ggplot(aes(fill = fct_infreq(extensiveContact), x = relYear.cut)) +
  geom_bar(position = "dodge", width = .5) + 
  facet_wrap(~`class/type`) +
  # geom_text(stat = 'count', 
  #           aes(label = paste0(after_stat(round(100*count/sum(count),1)), "%")), position = position_dodge(width = .9), hjust = -0.2) +
  labs(x = "Release Year", y = "Frequency", fill = "Extensive contact/feedback") +
  #scale_y_continuous(n.breaks = 13) + 
  scale_fill_brewer(palette = "Set1", direction = -1) +
  theme_light() +
  theme(strip.background =element_rect(fill="steelblue3"),
        strip.text = element_text(colour = 'white'),
        legend.position = "top") + 
  coord_flip()
```

```{r}
wdat %>% 
  mutate(softHealthMetrics.bi = if_else(softHealthMetrics == "No", true = "softHealthMetrics - No", false = "softHealthMetrics - Yes")) %>% 
  filter(!is.na(releaseYear)) %>% 
  mutate(relYear.cut = cut_interval(releaseYear, 4)) %>% 
  ggplot(aes(fill = fct_infreq(extensiveContact), x = relYear.cut)) +
  geom_bar(position = "dodge", width = .5) + 
  facet_wrap(~softHealthMetrics.bi, nrow = 2) +
  labs(x = "Release Year", y = "Frequency", fill = "Extensive contact/feedback") +
  # scale_y_continuous(n.breaks = 13) + 
  scale_fill_brewer(palette = "Set1", direction = 1) +
  theme_light() +
  theme(strip.background =element_rect(fill="steelblue3"),
        strip.text = element_text(colour = 'white'),
        legend.position = "top", panel.grid = element_blank() ) 
```

## Additional Plot: Diversity of the type of tools
```{r}
wdat %>% select(toolName = `tool name`, toolType = `class/type`) %>% 
  ggplot(aes(y = fct_infreq(toolType), fill = toolType)) +
  geom_bar(width = .5, show.legend = FALSE) + 
  labs(x = "Frequency", y = "Type of tool") +
  scale_x_continuous(n.breaks = 7) + 
  #scale_fill_brewer(palette = "Set3", direction = 1) +
  theme_light() +
  theme(panel.grid = element_blank() ) 
```



## Full dataset view 

```{r}
wdat %>% select(-`Instructions how to cite`, -`...11`) %>%  ## deleting these two columns to make the table more readable
  datatable(options = list(scrollX = TRUE))
```














